\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-09-30 gwnnDraft02.tex
A lot of grammar and consistency fixing.

The ``exotic sectors'' with $i<0$ provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 

@@

In general, the multi-hidden sector models explored feature a huge number of (nearly) massless degrees of freedom. Dark photons and dark neutrinos abound in these sectors and, assuming a relatively high reheat temperature, the leptons, quarks, and heavy bosons of these sectors can also be relativistic. In $N$naturalness this feature is realized quite dramatically: each of the $N$ of sectors possess relativistic degrees of freedom. The presence of these particles can have two main effects: extra relativistic particles can alter the expansion history of the universe through changes to the energy density or hidden sectors can feature annihilations that reheat the photons or neutrinos of our sector near Big Bang Nucleosynthesis (BBN) and affect the light element abundances. The effective number of neutrino species, $N_{eff}$, is impacted by these contributions and, as such, is the strictest constraint that must be dealt with when studying these type of multi-phase transition models. 

@@

for any \draft{check} decoupled hidden sector~\cite{Arkani-Hamed:2016rle}. Because the constraints on $N_{eff}$ are stronger at photon decoupling than at BBN, we can focus purely on the constraints provided by the former.

@@

with $T_{RH_{E1}}/T_{RH}$ being the ratio of the reheat temperatures of the 1st exotic sector and our own sector ($0.06$ in standard $N$naturalness with $r = 1$).  This sum is dominated by $i=1$, the sector with the lowest Higgs mass (and thus the most energy density) gives us a contribution of $\mathcal{O} (10^{-4})$. Evolving the sectors thermal histories forward in time to the recombination era gives us a slightly larger value, but still of order $\mathcal{O}(10^{-4})$; well under current CMB bounds. 

@@

We now turn to the gravitational wave signatures of our setup. At high temperatures, each of the hidden sectors has QCD in the quark/gluon phase, but at temperatures around $\Lambda_{\rm QCD,i}$, the sector undergoes a phase transition into the hadronic phase, where we compute $\Lambda_{\rm QCD,i}$ for the different sectors in Sec.~\ref{sec:dQCD}. As discussed in that section, this phase transition will be strongly first order (SFOPT) for certain numbers of light quarks, which will generate gravitational waves. This differs from QCD in the SM sector, as the PT is a crossover and not first order~\cite{Fodor:2001pe}. 

@@

where the time component has been integrated out due to nucleation occuring, not in vaccum, but in a finite temperature plasma. $\phi$ is the symmetry breaking scalar field with a non-zero VEV. In the case of the chiral phase transition, the scalar field breaking the SU$(N_f)_{R}$ $ \times$ SU$(N_f)_{L}$ chiral symmetry is the effective quark condensate $\phi_{i} \sim \langle q\bar{q} \rangle_{i}$ of the respective sector. We leave the thermalized potential $V(\phi, T)$ general. As previously stated, an exact QCD potential at the time of the chiral phase transition is not well understood outside of lattice results. In~\cite{Bai:2018dxf} chiral effective Lagrangian was used to calculate a low energy thermalized potential for confining SU$(N)$. 
The amount of energy density dumped into the individual sectors dictates the energy budget for the PT and hence for the gravitational waves. Assuming that the SM sector is radiation dominated, a quantity that characterizes the strength of the PT is the ratio of the latent heat of the phase transition $\epsilon$ to the energy density of radiation, at the time of nucleation~\citep{Espinosa:2010hh} ,

@@

with $\epsilon$ being calculable from the scalar potential. Assuming that there is a negligible amount of energy being dumped back into the SM, which would cause significant reheating of $\rho_{\gamma}$,  the latent head, $\epsilon$ should correspond to the energy density of the hidden sector going through the PT. The parameter $g_{*}$ in the denominator of Eq.~(\ref{eq:alpha}) is the number of relativistic degrees of freedom at the time of the phase transition, with contributions from species in both the visible and dark sectors. It has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective temperature ratios (energy densities)

@@

\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: MTG shocks within the plasma, left over from the sound wave propagation, will produce gravitational waves. With efficiency factor $\kappa_{

@@

where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectfully. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,

@@

This is not the case if we consider the more generalized reheating scenarios of section~\ref{sec:constraints}. Once the restriction that small vevs are preferentially reheated has been lifted, we can explore a landscape of standard sectors with much heavier particles (specifically quarks) and, as shown in section~\ref{sec:dQCD}, the SFOPT these entail. The key phenomenological constraint on such sectors is $\Delta N_{eff}$, giving us a maximum allowed temperature ratio (when compared to the SM) for each reheated hidden sector: Eq.~(\ref{eqn:energyDensityAllowed}) shows the maximum temperature ratios for specific numbers of additional hidden sectors. There are four scenarios that we examine here:

@@

The GW results of these cases are presented in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. When relaxing the assumptions on $\beta/H$, the scenarios in figure~(\ref{fig:Haa}) are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$ The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached at which two (or multiple) peaks, if the amplitudes are comparable. 

>>>>

The ``exotic sectors'' with $i<0$ provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{Arkani-Hamed:2016rle}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 

@@

In general, the multi-hidden sector models explored feature a huge number of (nearly) massless degrees of freedom. Dark photons and dark neutrinos abound in these sectors and, assuming a relatively high reheat temperature, the leptons, quarks, and heavy bosons of these sectors can also be relativistic. In $N$naturalness this feature is realized quite dramatically: each of the $N$ sectors possess relativistic degrees of freedom. The presence of these particles can have two main effects: extra relativistic particles can alter the expansion history of the universe through changes to the energy density or hidden sectors can feature annihilations that reheat the photons or neutrinos of our sector near Big Bang Nucleosynthesis (BBN) and affect the light element abundances. The effective number of neutrino species, $N_{eff}$, is impacted by these contributions and, as such, is the strictest constraint that must be dealt with when studying these type of multi-phase transition models. 

@@

for any decoupled hidden sector~\cite{Arkani-Hamed:2016rle}. Because the constraints on $N_{eff}$ are stronger at photon decoupling than at BBN, we can focus purely on the constraints provided by the former.

@@

with $T_{RH_{E1}}/T_{RH}$ being the ratio of the reheat temperatures of the 1st exotic sector and our own sector ($0.06$ in standard $N$naturalness with $r = 1$).  This sum is dominated by $i=1$, the sector with the lowest Higgs mass (and thus the most energy density) gives us a contribution of $\mathcal{O} (10^{-4})$ to $\Delta N_{eff}$. Evolving the sectors thermal histories forward in time to the recombination era gives us a slightly larger value, but still of order $\mathcal{O}(10^{-4})$, well under current CMB bounds. 

@@

We now turn to the gravitational wave signatures of our setup. At high temperatures, each of the hidden sectors has QCD in the quark/gluon phase, but at temperatures around $\Lambda_{\rm QCD,i}$, the $i^{\mathrm{th}}$ sector undergoes a phase transition into the hadronic phase that we computed for the different sectors in Sec.~\ref{sec:dQCD}. As discussed in that section, this phase transition will be strongly first order (SFOPT) for certain numbers of light quarks, which will generate gravitational waves. This differs from QCD in the SM sector, as the PT is a crossover and not first order~\cite{Fodor:2001pe}. 

@@

where the time component has been integrated out due to nucleation occuring not in vaccum but in a finite temperature plasma. $\phi$ is the symmetry breaking scalar field with a non-zero vev. In the case of the chiral phase transition, the scalar field breaking the SU$(N_f)_{R}$ $ \times$ SU$(N_f)_{L}$ chiral symmetry is the effective quark condensate $\phi_{i} \sim \langle q\bar{q} \rangle_{i}$ of the respective sector. We leave the thermalized potential $V(\phi, T)$ general. As previously stated, an exact QCD potential at the time of the chiral phase transition is not well understood outside of lattice results. In~\cite{Bai:2018dxf} chiral effective Lagrangian was used to calculate a low energy thermalized potential for confining SU$(N)$. 
The amount of energy density dumped into the individual sectors dictates the energy budget for the PT and hence for the gravitational waves. Assuming that the SM sector is radiation dominated, a quantity that characterizes the strength of the PT is the ratio of the latent heat of the phase transition, $\epsilon$, to the energy density of radiation, at the time of nucleation~\citep{Espinosa:2010hh} ,

@@

with $\epsilon$ being calculable from the scalar potential. Assuming that there is a negligible amount of energy being dumped back into the SM, which would cause significant reheating of $\rho_{\gamma}$,  the latent heat, $\epsilon$ should correspond to the energy density of the hidden sector going through the PT. The parameter $g_{*}$ in the denominator of Eq.~(\ref{eq:alpha}) is the number of relativistic degrees of freedom at the time of the phase transition, with contributions from species in both the visible and dark sectors. It has weak temperature dependence in a single sector, but when dealing with multiple hidden sectors, $g_{*}$ gains contributions from all $N$ sector's relativistic degrees of freedom, weighted by their respective temperature ratios (energy densities)

@@

\item \textbf{Magnetohydrodynamical contributions $\Omega_{B}$}: Turbulence within the plasma, left over from the sound wave propagation, will produce gravitational waves. With efficiency factor $\kappa_{

@@

where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,

@@

This is not the case if we consider the more generalized reheating scenarios of Sec.~\ref{sec:constraints}. Once the restriction that small vevs are preferentially reheated has been lifted, we can explore a landscape of standard sectors with much heavier particles (specifically quarks) and, as shown in section~\ref{sec:dQCD}, the SFOPT these entail. The key phenomenological constraint on such sectors is $\Delta N_{eff}$, giving us a maximum allowed temperature ratio (when compared to the SM) for each reheated hidden sector: Eq.~(\ref{eqn:energyDensityAllowed}) shows the maximum temperature ratios for specific numbers of additional hidden sectors. There are four scenarios that we examine here:

@@

The GW results of these cases are presented in figure~(\ref{fig:Haa}). In all cases, the summed GW signal is detectable by one or more proposed interferometers. When relaxing the assumptions on $\beta/H$, the scenarios in figure~(\ref{fig:Haa}) are still detectable for values ranging between $\mathcal{O}(1)$ and $\mathcal{O}(100)$ The frequency dependence in Eq.~(\ref{eqn::spectralshape}) takes the form of $f/f_{p}$, this causes a cancellation between the redshifting factors. As multiple sectors phase transition at different times, and therefore temperatures, the peaks will shift relative to each other, purely from the linear temperature dependence of the peak frequency $f_{p} \sim T_{\gamma}$.  This is seen in figure~(\ref{fig:Haa}), where the spectrum peaks are shifted causing a peak broadening of the convoluted spectrum. The broadening can be substantial if the hidden sectors transition between a large gap of time (temperature). Eventually, a temperature limit will be reached where two (or multiple) distinct peaks will be visible, provided that the amplitudes are comparable. 
